Geospatial - Assignment 3

### 2025-02-13 hw3/hw3_MoritzPeistNoemiLucchiSimonVellin.Rmd
Added pictures for comparisons
- Ignoring large geospatial files now - Rendering graphics in SVG and fixed notebook

```{r setup, include=FALSE}
# Disable warning messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

@@

```{r}

# Import SPEI index 
spei_index <- rast("/hw3/data/spei01.nc")

# Import US regions
us_states <- spData::us_states

```

@@

```{r}

# Get dates from SPEI dataset
dates <- time(spei_index)

@@

# Zonal statistics to compute the mean (each row is a country and each column is a date)
zonal_stats <- exact_extract(
  spei_subset,
  us_states,
  fun = "mean")

# Convert the wide format panel to long format panel and add state/region identifiers
# This is necessary to compute the yearly mean per country and then the yearly mean per region
spei_long <- zonal_stats %>%
  as.data.frame() %>%
  # Convert from wide to long format
  pivot_longer(
    cols = everything(),
    names_to = "time_index",
    values_to = "spei") %>%
  # Add state and region identifiers, plus temporal information
  mutate(
    state = rep(us_states$NAME, each = length(dates)), 
    region = rep(us_states$REGION, each = length(dates)),  
    date = rep(dates, times = nrow(us_states)),  
    year = year(date),
    month = month(date)) %>%
  select(state, region, year, month, spei)

# Compute the yearly mean for each state 
spei_state_annual <- spei_long %>%
  group_by(state, region, year) %>%
  summarise(
    mean_spei_state = mean(spei, na.rm = TRUE),  
    .groups = "drop")

# Compute the yearly mean for each region 
spei_regional <- spei_state_annual %>%
  group_by(region, year) %>%
  summarise(
    mean_spei_region = mean(mean_spei_state, na.rm = TRUE),
    .groups = "drop")

```

@@

```{r}

# Plotting
ggplot(spei_regional, aes(x = year, y = mean_spei_region, color = region, group = region)) +
  geom_line(linewidth = 0.8) +  
  geom_smooth(data = spei_regional, aes(x = year, y = mean_spei_region), 
              method = "loess", color = "blue", se = TRUE, linewidth = 1.2, inherit.aes = FALSE) +
  scale_color_brewer(palette = "Set2") +
  scale_x_continuous(
    breaks = seq(1970, 2010, by = 10),
    expand = c(0.02, 0.02)) +
  scale_y_continuous(
    limits = c(-1, 1),  
    breaks = seq(-1, 1, by = 1)) +
  labs(
    title = "SPEI Index by US Region (1965-2015)",
    x = "Year",
    y = "SPEI Index",
    color = "Region")

```

@@

>>>>

```{r setup, include=FALSE}
# Load required package for SVG
library(ragg)

# Disable warning messages and SVG plots
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  # fig.ext = "svg",
  fig.width = 16,
  fig.height = 8,
  fig.retina = 2,
  dpi = 300,
  dev = "svg"
)
```

@@

```{r}
# Import SPEI index
spei_index <- rast("data/ex1/spei01.nc")

# Import US regions
us_states <- spData::us_states
```

@@

```{r}
# Get dates from SPEI dataset
dates <- time(spei_index)

@@

# Zonal statistics to compute the mean (each row is a country and each column is a date)
zonal_stats <- exact_extract(
  spei_subset,
  us_states,
  fun = "mean"
)

# Convert the wide format panel to long format panel and add state/region identifiers
# This is necessary to compute the yearly mean per country and then the yearly mean per region
spei_long <- zonal_stats %>%
  as.data.frame() %>%
  # Convert from wide to long format
  pivot_longer(
    cols = everything(),
    names_to = "time_index",
    values_to = "spei"
  ) %>%
  # Add state and region identifiers, plus temporal information
  mutate(
    state = rep(us_states$NAME, each = length(dates)),
    region = rep(us_states$REGION, each = length(dates)),
    date = rep(dates, times = nrow(us_states)),
    year = year(date),
    month = month(date)
  ) %>%
  select(state, region, year, month, spei)

# Compute the yearly mean for each state
spei_state_annual <- spei_long %>%
  group_by(state, region, year) %>%
  summarise(
    mean_spei_state = mean(spei, na.rm = TRUE),
    .groups = "drop"
  )

# Compute the yearly mean for each region
spei_regional <- spei_state_annual %>%
  group_by(region, year) %>%
  summarise(
    mean_spei_region = mean(mean_spei_state, na.rm = TRUE),
    .groups = "drop"
  )
```

@@

```{r}
# Plotting
ggplot(spei_regional, aes(x = year, y = mean_spei_region, color = region, group = region)) +
  geom_line(linewidth = 0.8) +
  geom_smooth(
    data = spei_regional, aes(x = year, y = mean_spei_region),
    method = "loess", color = "blue", se = TRUE, linewidth = 1.2, inherit.aes = FALSE
  ) +
  scale_color_brewer(palette = "Set2") +
  scale_x_continuous(
    breaks = seq(1970, 2010, by = 10),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 1)
  ) +
  labs(
    title = "SPEI Index by US Region (1965-2015)",
    x = "Year",
    y = "SPEI Index",
    color = "Region"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
  )
```

In comparison the first graph looked as follows:

![Figure 1](data/ex1/raster09.png)

@@

...

In comparison the first map looked as follows:

![Map 1](data/ex2/raster07.png)

...

and the second graph as follows:

![Figure 2](data/ex2/raster08.png)
